---
title: "BST210_DataCleaning"
date: "2022-10-04"
output: html_document
---

```{r pressure, echo=FALSE}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library("haven")
library(dplyr)
library(tidyr)
library(broom)
library(biglm)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
dat_asthma = read.csv("CDI_Asthma.csv")
dat_smk = read.csv("CDI_Tobacco.csv")
dat_alc = read.csv("CDI_Alcohol.csv")
dat_obe = read.csv("CDI_Obesity.csv")
dat_prox = read.csv("CDI_Overarching.csv")
```

#### Choosing target group
Choose target group to be "women aged 18-44 years" for age and gender restricted data
Crude prevalence
```{r}
#clean asthma data to restrict target group
dat_asthma <- dat_asthma[dat_asthma$QuestionID == "AST1_2",]
#drop missing values for crude rate
dat_asthma <- dat_asthma[complete.cases(dat_asthma[,c("DataValue")]),]
#sort dataframe by year
dat_asthma <- dat_asthma[order(dat_asthma$YearStart, decreasing = FALSE),]
```

Tobacco data: Current cigarette smoking among women aged 18-44 years
crude prevalence
```{r}
#clean Tobacco data
dat_smk <- dat_smk[dat_smk$QuestionID == "TOB1_3",]
dat_smk <- dat_smk[complete.cases(dat_smk[,c("DataValue")]),]
dat_smk <- dat_smk[order(dat_smk$YearStart, decreasing = FALSE),]
```

Alcohol data: 
Heavy drinking among women aged 18-44 years: QuestionID = ALC5_2
Binge drinking prevalence among women aged 18-44 years: QuestionID = ALC2_3
```{r}
dat_alc <- filter(dat_alc, dat_alc$QuestionID %in% c("ALC5_2", "ALC2_3"))
dat_alc <- dat_alc[complete.cases(dat_alc[,c("DataValue")]),]
dat_alc <- dat_alc[order(dat_alc$YearStart, decreasing = FALSE),]
```

Obesity data: Overweight or obesity among women aged 18-44 years
Crude prevalence
```{r}
dat_obe <- dat_obe[dat_obe$QuestionID == "NPAW2_3",]
dat_obe <- dat_obe[complete.cases(dat_obe[,c("DataValue")]),]
dat_obe <- dat_obe[order(dat_obe$YearStart, decreasing = FALSE),]
```
 
Overarching data: 
High school completion among women aged 18-44 years: OVC2_2
Self-rated health status among women aged 18-44 years: OVC6_2
Current health care coverage among women aged 18-44 years: OVC1_2
```{r}
dat_prox <- filter(dat_prox, dat_prox$QuestionID %in% c("OVC2_2", "OVC6_2","OVC1_2"))
dat_prox <- dat_prox[complete.cases(dat_prox[,c("DataValue")]),]
dat_prox <- dat_prox[order(dat_prox$YearStart, decreasing = FALSE),]
```



