---
title: "BST210_DataCleaning"
date: "2022-10-04"
output: html_document
---

```{r pressure, echo=FALSE}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library(dplyr)
library(tidyr)
library(broom)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
#<<<<<<< HEAD
library(data.table)
dat_asthma = read.csv("CDI_Asthma.csv")
dat_smk = read.csv("CDI_Tobacco.csv")
dat_alc = read.csv("CDI_Alcohol.csv")
dat_obe = read.csv("CDI_Obesity.csv")
dat_prox = read.csv("CDI_Overarching.csv")
#--------------=======
#>>>>>>> aac3615087f90365cebdcffeaa02ac77b7ef9bc9
```

#### Choosing target group
Choose target group to be "women aged 18-44 years" for age and gender restricted data
Crude prevalence
```{r}
#clean asthma data to restrict target group
dat_asthma <- read.csv("CDI_Asthma.csv") |> filter(QuestionID=="AST1_2")
#drop missing values for crude rate
#####dat_asthma <- drop_na(dat_asthma, DataValue)

#sort dataframe by year
dat_asthma <- dat_asthma[order(dat_asthma$YearStart, decreasing = FALSE),]
```

Tobacco data: Current cigarette smoking among women aged 18-44 years
crude prevalence
```{r}
#clean Tobacco data
dat_smk <- read.csv("CDI_Tobacco.csv")|> filter(QuestionID == "TOB1_3")
dat_smk <- dat_smk[order(dat_smk$YearStart, decreasing = FALSE),]
```

Alcohol data: 
Heavy drinking among women aged 18-44 years: QuestionID = ALC5_2
Binge drinking prevalence among women aged 18-44 years: QuestionID = ALC2_3
```{r}
dat_heavyDrink <- read.csv("CDI_Alcohol.csv") |> filter(QuestionID == "ALC5_2")
dat_bingeDrink <- read.csv("CDI_Alcohol.csv") |> filter(QuestionID == "ALC2_3")
#####dat_alc <- dat_alc[order(dat_alc$YearStart, decreasing = FALSE),]
```

Obesity data: Overweight or obesity among women aged 18-44 years
Crude prevalence
```{r}
dat_obe <- read.csv("CDI_Obesity.csv")|> filter(QuestionID == "NPAW2_3")
#dat_obe <- dat_obe[order(dat_obe$YearStart, decreasing = FALSE),]
```
 
Overarching data: 
High school completion among women aged 18-44 years: OVC2_2
Self-rated health status among women aged 18-44 years: OVC6_2
Current health care coverage among women aged 18-44 years: OVC1_2
```{r}
#high school data no stratification
dat_highSch <- read.csv("CDI_Overarching.csv")|> filter(QuestionID == "OVC2_2")
dat_selfHlthSta <- read.csv("CDI_Overarching.csv")|> filter(QuestionID == "OVC6_2")
dat_HlthCare <- read.csv("CDI_Overarching.csv")|> filter(QuestionID == "OVC1_2")
#####dat_prox <- dat_prox[order(dat_prox$YearStart, decreasing = FALSE),]
```

Clean Daily Summary data to obtain annual average by state from 2011 to 2019 (Ozone, SO2, CO, NO2)
```{r}
#exposure variables defined as percentage of AQI days with co, no2, o3, pm2.5, pm10 as primary pollutants
# naming pattern: Ozone (44201)	SO2 (42401)	CO (42101)	NO2 (42602) PM2.5 FRM/FEM Mass (88101)

#Ozone unit: parts per million--------------------
#read in daily summary file - O3
file_o3 <- list.files(path = "./exposure_data", pattern="daily_44201", full.names =T)
dat_o3_tibble <- lapply(file_o3, read.csv)

#read in files of land area of U.S. counties
# LND110210D Land area in square miles 2010
land_area <- readxl::read_xls("LND01.xls") %>% select(Areaname, STCOU, LND110210D) %>% 
  mutate(State.Code = substr(STCOU, start = 1, stop = 2),
         County.Code = substr(STCOU, start = 3, stop = 5)) %>% subset(County.Code != "000") %>% 
  rename(land.area = LND110210D) %>% select(State.Code, County.Code, land.area)

land_area$State.Code <-  gsub("(?<![0-9])0+", "", land_area$State.Code, perl = TRUE) %>% as.integer()
land_area$County.Code <- gsub("(?<![0-9])0+", "", land_area$County.Code, perl = TRUE) %>% as.integer()


#create function to find the annual average of ozone exposure by state
find.state.annual.o3 <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(o3 = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(o3_mean = weighted.mean(o3, land.area))

  return(annual.mean)
}

dat_o3_tibble_mean <- lapply(dat_o3_tibble, find.state.annual.o3)

for (i in 1:length(dat_o3_tibble_mean)){
  dat_o3_tibble_mean[[i]] <- dat_o3_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_o3 <- rbindlist(dat_o3_tibble_mean, use.names = TRUE, fill = TRUE)


#pm2.5 unit: microgram/cubic meter-----------------
file_pm2.5 <- list.files(path = "./exposure_data", pattern="daily_88101", full.names =T)
<<<<<<< HEAD
```

```{r}
# data cleaning for socio economic factors
homeless <-  read_csv('homelessness.csv') 
snap<- read_csv('SNAPenrollment.csv')
unrate <- read_csv('unemploymentrate.csv')
poverty <- read.csv('poverty_gender.csv')

```

```{r}
states_vector <- c(unique(homeless$Location), unique(poverty$Location), unique(snap$Location), unique(unrate$State))
states_vector

unique_states <- intersect(unique(homeless$Location), unique(poverty$Location))
unique_states<- intersect(unique_states, unique(snap$Location))
unique_states <- intersect(unique_states, unique(unrate$State))

```



```{r}
names(unrate)[names(unrate) == 'State'] <- 'Location'
```


```{r}
one<-merge(homeless, unrate, by = c('Location', 'year'))
two <- merge(snap, one, by= c('Location', 'year'))
soecon <- merge(two, poverty, by = c('Location', 'year'))
names(soecon)[names(soecon) == 'Individuals'] <- 'individual_homeless'
names(soecon)[names(soecon) == 'Total People Experiencing Homelessness'] <- 'total_homeless'
names(soecon)[names(soecon) == 'People in Families with Children'] <- 'fam_homeless'

```


```{r}
unique(soecon$Location)
```

unique(soecon$Location)

```{r}
dat_pm2.5_tibble <- lapply(file_pm2.5, read.csv)

#read in files of land area of U.S. counties
# LND110210D Land area in square miles 2010
land_area <- readxl::read_xls("LND01.xls") %>% select(Areaname, STCOU, LND110210D) %>% 
  mutate(State.Code = substr(STCOU, start = 1, stop = 2),
         County.Code = substr(STCOU, start = 3, stop = 5)) %>% subset(County.Code != "000") %>% 
  rename(land.area = LND110210D) %>% select(State.Code, County.Code, land.area)

land_area$State.Code <-  gsub("(?<![0-9])0+", "", land_area$State.Code, perl = TRUE) %>% as.integer()
land_area$County.Code <- gsub("(?<![0-9])0+", "", land_area$County.Code, perl = TRUE) %>% as.integer()


#create function to find the annual average of pm2.5 exposure by state
find.state.annual.pm2.5 <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(pm2.5 = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(pm2.5_mean = weighted.mean(pm2.5, land.area))

  return(annual.mean)
}

dat_pm2.5_tibble_mean <- lapply(dat_pm2.5_tibble, find.state.annual.pm2.5)

for (i in 1:length(dat_pm2.5_tibble_mean)){
  dat_pm2.5_tibble_mean[[i]] <- dat_pm2.5_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_pm2.5 <- rbindlist(dat_pm2.5_tibble_mean, use.names = TRUE, fill = TRUE)

#SO2 unit: parts per billion-----------------
file_so2 <- list.files(path = "./exposure_data", pattern="daily_42401", full.names =T)

dat_so2_tibble <- lapply(file_so2, read.csv)

#create function to find the annual average of pm2.5 exposure by state
find.state.annual.so2 <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(so2 = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(so2_mean = weighted.mean(so2, land.area))

  return(annual.mean)
}

dat_so2_tibble_mean <- lapply(dat_so2_tibble, find.state.annual.so2)

for (i in 1:length(dat_so2_tibble_mean)){
  dat_so2_tibble_mean[[i]] <- dat_so2_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_so2 <- rbindlist(dat_so2_tibble_mean, use.names = TRUE, fill = TRUE)

#CO unit: parts per million-----------------
file_co <- list.files(path = "./exposure_data", pattern="daily_42101", full.names =T)

dat_co_tibble <- lapply(file_co, read.csv)

#create function to find the annual average of pm2.5 exposure by state
find.state.annual.co <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(co = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(co_mean = weighted.mean(co, land.area))

  return(annual.mean)
}

dat_co_tibble_mean <- lapply(dat_co_tibble, find.state.annual.co)

for (i in 1:length(dat_co_tibble_mean)){
  dat_co_tibble_mean[[i]] <- dat_co_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_co <- rbindlist(dat_co_tibble_mean, use.names = TRUE, fill = TRUE)





#NO2 unit: parts per billion --------------------
#read in daily summary file - no2
file_no2 <- list.files(path = "./exposure_data", pattern="daily_42602", full.names =T)
dat_no2_tibble <- lapply(file_no2, read.csv)

#create function to find the annual average of ozone exposure by state
find.state.annual.no2 <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(no2 = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(no2_mean = weighted.mean(no2, land.area))

  return(annual.mean)
}

dat_no2_tibble_mean <- lapply(dat_no2_tibble, find.state.annual.no2)

for (i in 1:length(dat_no2_tibble_mean)){
  dat_no2_tibble_mean[[i]] <- dat_no2_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_no2 <- rbindlist(dat_no2_tibble_mean, use.names = TRUE, fill = TRUE)


dat_exp <- dat_o3 %>% left_join(dat_pm2.5, by = c("State.Code","year")) %>% left_join(dat_so2, by = c("State.Code","year")) %>% 
  left_join(dat_co, by = c("State.Code","year")) %>% left_join(dat_no2, by = c("State.Code","year"))
````
