---
title: "BST210_DataCleaning"
date: "2022-10-04"
output: html_document
---

```{r pressure, echo=FALSE}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library("haven")
library(dplyr)
library(tidyr)
library(broom)
library(biglm)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(data.table)
dat_asthma = read.csv("CDI_Asthma.csv")
dat_smk = read.csv("CDI_Tobacco.csv")
dat_alc = read.csv("CDI_Alcohol.csv")
dat_obe = read.csv("CDI_Obesity.csv")
dat_prox = read.csv("CDI_Overarching.csv")
```

#### Choosing target group
Choose target group to be "women aged 18-44 years" for age and gender restricted data
Crude prevalence
```{r}
#clean asthma data to restrict target group
dat_asthma <- dat_asthma[dat_asthma$QuestionID == "AST1_2",]
#drop missing values for crude rate
dat_asthma <- dat_asthma[complete.cases(dat_asthma[,c("DataValue")]),]
#sort dataframe by year
dat_asthma <- dat_asthma[order(dat_asthma$YearStart, decreasing = FALSE),]
```

Tobacco data: Current cigarette smoking among women aged 18-44 years
crude prevalence
```{r}
#clean Tobacco data
dat_smk <- dat_smk[dat_smk$QuestionID == "TOB1_3",]
dat_smk <- dat_smk[complete.cases(dat_smk[,c("DataValue")]),]
dat_smk <- dat_smk[order(dat_smk$YearStart, decreasing = FALSE),]
```

Alcohol data: 
Heavy drinking among women aged 18-44 years: QuestionID = ALC5_2
Binge drinking prevalence among women aged 18-44 years: QuestionID = ALC2_3
```{r}
dat_alc <- filter(dat_alc, dat_alc$QuestionID %in% c("ALC5_2", "ALC2_3"))
dat_alc <- dat_alc[complete.cases(dat_alc[,c("DataValue")]),]
dat_alc <- dat_alc[order(dat_alc$YearStart, decreasing = FALSE),]
```

Obesity data: Overweight or obesity among women aged 18-44 years
Crude prevalence
```{r}
dat_obe <- dat_obe[dat_obe$QuestionID == "NPAW2_3",]
dat_obe <- dat_obe[complete.cases(dat_obe[,c("DataValue")]),]
dat_obe <- dat_obe[order(dat_obe$YearStart, decreasing = FALSE),]
```
 
Overarching data: 
High school completion among women aged 18-44 years: OVC2_2
Self-rated health status among women aged 18-44 years: OVC6_2
Current health care coverage among women aged 18-44 years: OVC1_2
```{r}
dat_prox <- filter(dat_prox, dat_prox$QuestionID %in% c("OVC2_2", "OVC6_2","OVC1_2"))
dat_prox <- dat_prox[complete.cases(dat_prox[,c("DataValue")]),]
dat_prox <- dat_prox[order(dat_prox$YearStart, decreasing = FALSE),]
```

Clean Daily Summary data to obtain annual average by state from 2011 to 2019 (Ozone, SO2, CO, NO2)
```{r}
#exposure variables defined as percentage of AQI days with co, no2, o3, pm2.5, pm10 as primary pollutants
# naming pattern: Ozone (44201)	SO2 (42401)	CO (42101)	NO2 (42602) PM2.5 FRM/FEM Mass (88101)


#read in daily summary file - O3
file_o3 <- list.files(path = "./exposure_data", pattern="daily_44201", full.names =T)
dat_o3_tibble <- lapply(file_o3, read.csv)

#read in files of land area of U.S. counties
# LND110210D Land area in square miles 2010
land_area <- readxl::read_xls("LND01.xls") %>% select(Areaname, STCOU, LND110210D) %>% 
  mutate(State.Code = substr(STCOU, start = 1, stop = 2),
         County.Code = substr(STCOU, start = 3, stop = 5)) %>% subset(County.Code != "000") %>% 
  rename(land.area = LND110210D) %>% select(State.Code, County.Code, land.area)

land_area$State.Code <-  gsub("(?<![0-9])0+", "", land_area$State.Code, perl = TRUE) %>% as.integer()
land_area$County.Code <- gsub("(?<![0-9])0+", "", land_area$County.Code, perl = TRUE) %>% as.integer()


#create function to find the annual average of ozone exposure by state
find.state.annual.o3 <- function(x) {
  annual.mean <- x %>% group_by(State.Code, County.Code) %>% mutate(o3 = mean(Arithmetic.Mean)) %>% 
    left_join(land_area, by = c("State.Code", "County.Code")) %>% ungroup() %>% 
    group_by(State.Code) %>% summarise(o3_mean = weighted.mean(o3, land.area))

  return(annual.mean)
}

dat_o3_tibble_mean <- lapply(dat_o3_tibble, find.state.annual.o3)

for (i in 1:length(dat_o3_tibble_mean)){
  dat_o3_tibble_mean[[i]] <- dat_o3_tibble_mean[[i]] %>% mutate(year = i + 2010)
}

dat_o3 <- rbindlist(dat_o3_tibble_mean, use.names = TRUE, fill = TRUE)


file_pm2.5 <- list.files(path = "./exposure_data", pattern="daily_88101", full.names =T)
````
